---
title: "Bellabeat Case Study"
author: "Lidiya Arsova"
date: "2025-04-13"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE)
library(tidyverse)
library(lubridate)
library(ggplot2)
```

## Ask

Bellabeat is a high-tech wellness company that manufactures health-focused smart devices for women. The goal of this project is to analyze Fitbit data and uncover user trends that Bellabeat can use to inform its marketing strategy.

## Prepare

We are working with public Fitbit data collected from 30 users over two months, shared under a public domain license via Kaggle and Zenodo.

We are using two daily activity files — one from each month — and will combine them.

```{r}
daily1 <- read_csv("dailyActivity_merged.csv")
daily2 <- read_csv("dailyActivity_merged2.csv")
```

## Process

Here, we combine the two datasets, remove any duplicates, and convert the activity date column into proper Date format.

```{r}
daily_full <- bind_rows(daily1, daily2) %>%
  distinct() %>%
  mutate(ActivityDate = mdy(ActivityDate))
```
## Analyze

We calculate summary statistics to understand users' daily activity patterns.

```{r}
daily_full %>%
  summarise(
    avg_steps = mean(TotalSteps, na.rm = TRUE),
    avg_calories = mean(Calories, na.rm = TRUE),
    avg_very_active = mean(VeryActiveMinutes, na.rm = TRUE),
    avg_lightly_active = mean(LightlyActiveMinutes, na.rm = TRUE),
    avg_sedentary = mean(SedentaryMinutes, na.rm = TRUE)
  )
```

We also explore how different activities correlate with calories burned.

```{r}
cor_data <- daily_full %>%
  select(TotalSteps, TotalDistance, VeryActiveMinutes, FairlyActiveMinutes,
         LightlyActiveMinutes, SedentaryMinutes, Calories)

cor(cor_data, use = "complete.obs")
```

## Share

### Average Calories Burned by Activity Level

```{r}
daily_full %>%
  pivot_longer(cols = c(VeryActiveMinutes, FairlyActiveMinutes, LightlyActiveMinutes),
               names_to = "ActivityLevel", values_to = "Minutes") %>%
  group_by(ActivityLevel) %>%
  summarise(AvgCalories = mean(Calories, na.rm = TRUE)) %>%
  ggplot(aes(x = ActivityLevel, y = AvgCalories, fill = ActivityLevel)) +
  geom_col() +
  labs(title = "Avg Calories Burned by Activity Level", y = "Calories", x = "Activity Level") +
  theme_minimal()
```

### Average Steps by Weekday

```{r}
daily_full %>%
  mutate(Weekday = wday(ActivityDate, label = TRUE)) %>%
  group_by(Weekday) %>%
  summarise(AverageSteps = mean(TotalSteps, na.rm = TRUE)) %>%
  ggplot(aes(x = Weekday, y = AverageSteps)) +
  geom_col(fill = "steelblue") +
  labs(title = "Average Steps by Weekday", x = "Day of Week", y = "Steps") +
  theme_minimal()
```
## Act

Based on this analysis, Bellabeat could take the following actions:

Encourage High-Intensity Movement:
Very Active Minutes had a strong correlation with calories burned. Bellabeat can promote workouts that drive this metric.

Target Weekday Engagement:
Most users are more active during weekdays. Bellabeat can use this insight to drive weekday-focused campaigns.

Combat Sedentary Behavior:
Sedentary minutes are still high. Bellabeat can add movement reminders or gentle nudge notifications via smart devices.

## Conclusion

This project explored user activity data to generate business insights for Bellabeat. Using R and tidyverse, we cleaned and analyzed the data, created visualizations, and delivered actionable recommendations for smarter engagement strategies.




